---
title: 'Lab 11: Simulation and Power Analysis'
author: "Mandy Klehr -- I worked with Matt, John, Heather, Jahiya, and Bonnie"
date: "12/16/2021"
output:  
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r walkthough setting up bird and habitat dataframe, message = FALSE}

require(here)

dat_bird = read.csv(here("data", "bird.sub.csv"))
dat_hab = read.csv(here("data", "hab.sub.csv"))
birdhab = merge(dat_bird, dat_hab)

dim(birdhab)

```

```{r walkthrough fit linear model to observed data, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}


fit_1 = lm(BRCR ~ ls, data = birdhab)

plot(x = birdhab$ls, y = birdhab$BRCR)
abline(fit_1)

summary(fit_1)

```

```{r walkthrough create linear function and linear simulator, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}

linear = function(x, y_int, slope)
  {
    y_int + slope * x
  }

#test implementation of linear function

linear(x = 1, y_int = 1, slope = 1)
linear(x = 3:5, y_int = 1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 0.01)

linear_simulator = function(x, y_int, slope, st_dev)
  {
  linear(x, y_int, slope) + rnorm(length(x), mean = 0, sd = st_dev)
  }

n = 200

par(mfrow = c(2, 2))
for (i in 1:4)
{
  x = runif(n = n)
  plot(
    x,
    linear_simulator(x, y_int = 1, slope = 4.5, st_dev = 0.1),
    main = "", xlab = "x", ylab = "y",
    pch = 16, col = rgb(0, 0.2, 0, 0.2))
}

n = 400

par(mfrow = c(2, 2))
for (i in 1:4)
{
  x = runif(n = n)
  plot(
    x, linear_simulator(x, y_int = 10, slope = -6.5, st_dev = 1.1),
    main = "", xlab = "x", ylab = "y",
    pch = 16, col = rgb(0, 0.2, 0, 0.2))
}

```
```{r walkthrough use observed data to simulate data, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}

#coefficients to extract intercept and slope values

fit_1_coefs = coefficients(fit_1)
str(fit_1_coefs)

#extract sd using summary and sigma

fit_1_summary = summary(fit_1)
str(fit_1_summary)
fit_1_summary$sigma

#store the intercept, slope, and sd from model into variables

int_obs = 0.0991
slope_obs = 0.00584
sd_obs = fit_1_summary$sigma

#plot the observed and simulated data

plot(
  x = birdhab$ls, 
  y = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  ),
  main = "Simulated Data",
  xlab = "late-successional forest",
  ylab = "Brown Creeper Abundance")

```


```{r walkthrough plot observed and simulated data, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}
#alternatively, plot observed first and ADD simulated data

plot(
  birdhab$ls, birdhab$BRCR, 
  xlab = "late-successional forest extent",
  ylab = "Brown Creeper abundance",
  pch = 19)

points(
  x = birdhab$ls, 
  y = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  ),
  col = adjustcolor("red", alpha = 0.3),
  pch = 16)

legend(
  "topleft",
  legend = c("data", "simulation"),
  pch = 16,
  col = c(1, adjustcolor("red", alpha = 0.3)))
```

```{r walkthrough power analysis for linear regression model, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}

y_sim = linear_simulator(
  x = birdhab$ls,
  y_int = int_obs,
  slope = slope_obs,
  st_dev = sd_obs
)

fit_sim = lm(y_sim ~ birdhab$ls)
summary(fit_sim)

sum_1 = summary(fit_sim)
sum_1$coefficients


#figure out the power

n_sims = 1000
p_vals = numeric(n_sims)
alpha = 0.05
for(i in 1:n_sims)
{
  y_sim = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  )
  fit_sim = lm(y_sim ~ birdhab$ls)
  
  p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
}
sum(p_vals < alpha) / n_sims


#function for linear simulation fits

linear_sim_fit = function(x, slope, y_int, st_dev)
{
  y_sim = linear_simulator(
    x = x,
    y_int = y_int,
    slope = slope,
    st_dev = st_dev
  )
  fit_sim = lm(y_sim ~ x)
  return(fit_sim)
}

#elaboration of the above code

alpha = 0.05
n_sims = 1000
p_vals = numeric(n_sims)

n_effect_sizes = 20
effect_sizes_1 = seq(-.01, .01, length.out = n_effect_sizes)

effect_size_powers = numeric(n_effect_sizes)

for(j in 1:n_effect_sizes)
{
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = birdhab$ls,
      y_int = int_obs,
      slope = effect_sizes_1[j],
      st_dev = sd_obs
    )
    
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  effect_size_powers[j] = sum(p_vals < alpha) / n_sims
}

sim_effect_size = 
  data.frame(
    power       = effect_size_powers,
    effect_size = effect_sizes_1)

#plot our simulation

plot(
  power ~ effect_size, data = sim_effect_size,
  type = 'l', xlab = 'Effect size', ylab = 'Power')
abline(v = slope_obs, lty = 2, col = 'red')


#do the same thing for a gradient of sample sizes

alpha = 0.05
n_sims = 1000
p_vals = numeric(n_sims)

sample_sizes = seq(5, 100)
sample_size_powers = numeric(length(sample_sizes))

for(j in 1:length(sample_sizes))
{
  x_vals = seq(0, 100, length.out = sample_sizes[j])
  
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = x_vals,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = sd_obs
    )
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  sample_size_powers[j] = sum(p_vals < alpha) / n_sims
}

sim_sample_size = 
  data.frame(
    power       = sample_size_powers,
    sample_size = sample_sizes)

plot(
  power ~ sample_size, data = sim_sample_size,
  type = 'l', xlab = 'Sample size', ylab = 'Power')
abline(v = nrow(birdhab), lty = 2, col = 'red')

```

```{r walkthrough power heat map bivariate power analysis, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}

alpha = 0.01
n_sims = 50

p_vals = numeric(n_sims)

n_effect_sizes = 20
effect_sizes = seq(-.01, .01, length.out = n_effect_sizes)

sample_sizes = seq(10, 50)

sim_output_2 = matrix(nrow = length(effect_sizes), ncol = length(sample_sizes))

for(k in 1:length(effect_sizes))
{
  effect_size = effect_sizes[k]
  for(j in 1:length(sample_sizes))
  {
    x_vals = seq(0, 100, length.out = sample_sizes[j])
    
    for(i in 1:n_sims)
    {
      fit_sim = linear_sim_fit(
        x = x_vals,
        y_int = int_obs,
        slope = effect_size,
        st_dev = sd_obs
      )
      p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
    }
    sim_output_2[k, j] = sum(p_vals < alpha) / n_sims
  }
  print(paste0("computing effect size ", k," of ", length(effect_sizes)))
}

sim_n_effect_size = 
  list(
    power = sim_output_2,
    effect_size = effect_sizes,
    sample_size = sample_sizes
  )

image(
  sim_n_effect_size$power,
  xlab = "Effect size",
  ylab = "Sample Size",
  axes = FALSE)

# add x-axis labels
axis(
  1, 
  at = c(0, 0.5, 1), 
  labels = c(-.01, 0.0, .01))

# add y=axis labels
axis(
  2, 
  at = c(0, 1), 
  labels = c(sample_sizes[1], tail(sample_sizes, 1)))
```

```{r contour plot, echo=FALSE, results = FALSE, message = FALSE, warning = FALSE, fig.show = 'hide'}
contour(
  x = sim_n_effect_size$effect_size,
  y = sim_n_effect_size$sample_size,
  z = sim_n_effect_size$power,
  xlab = "effect size",
  ylab = "sample size",
  main = "Contour Plot of Statistical Power",
  levels = seq(0, 1, length.out = 9),
  drawlabels = TRUE,
  # method = "simple")
  method = "edge")
```
```{r walkthrough perspective plot and 3d plots, echo=FALSE, fig.show= 'hide', message=FALSE, warning=FALSE, results=FALSE}

persp(
  x = sim_n_effect_size$effect_size,
  y = sim_n_effect_size$sample_size,
  z = sim_n_effect_size$power,
  xlab = "beta", ylab = "n", zlab = "power",
  col = 'lightblue',
  theta = 30, phi = 30, expand = .75,
  ticktype = 'detailed')


require(rgl)

#persp3d(
#  x = sim_n_effect_size$effect_size,
#  y = sim_n_effect_size$sample_size,
#  z = sim_n_effect_size$power,
#  xlab = "beta", ylab = "n", zlab = "power",
#  col = 'lightblue',
#  theta = 30, phi = 30, expand = .75,
#  ticktype = 'detailed')

#require(htmlwidgets)

#saveWidget(
#  rglwidget(),
#  file = here(
#    "docs", "webGL",
#    "n_effect_size_power_sim_plot.html"),
#  selfcontained = TRUE
#)


#save data to file when running large simulations

#save(
#  sim_n_effect_size,
#  file = here::here("data", "lab_11_n_effect_sizes.Rdata"))


#load data once it has been saves

#load(file = here::here("data", "lab_11_n_effect_sizes.Rdata")))


```

# Dispersion Simulation

*Create a line plot of dispersion (x-axis) and statistical power (y-axis) with a dotted vertical red line at the observed population standard deviation value.*

```{r }
alpha = 0.05
n_sims = 100
p_vals = numeric(n_sims)

# What was the observed standard deviation?
sd_obs

# specify the number of different standard deviation values to simulate:
n_sds = 20
pop_sds = seq(from = 0.01, to = 1.5, length.out = n_sds)

pop_sd_power = numeric(n_sds)

for(j in 1:length(pop_sds))
{
  pop_sd_j = pop_sds[j]
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(x = birdhab$ls,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = pop_sd_j)
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  pop_sd_power[j] = sum(p_vals < alpha) / n_sims
}

sim_output_dispersion = data.frame(
  sd = pop_sds,
  power = pop_sd_power)

# You should save your simulation results so you don't have to run it every time.
save(
  sim_output_dispersion, 
  file = here("data", "lab_ll_dat_dispersion_sim.RData"))

load(file = here("data", "lab_ll_dat_dispersion_sim.RData"))

```


# Q1

*Include a figure of your line plot in your report.*

```{r, echo = FALSE}
# Line plot of standard deviation (x-axis) and statistical power (y-axis)
plot(sim_output_dispersion,
  type = 'l', xlab = 'Standard deviation', ylab = 'Power')


# Add a dotted vertical red line at the observed population standard deviation value.
abline(v = sd_obs, lty = 2, col = 'red')

```


# Q2

*Why do you think that statistical power decreases as population dispersion increases?*

The statistical power decreases as population dispersion increases because dispersion decreases the probability of rejecting the null hypothesis (a **Type I error** when a relationship may actually exist) which may be indicative of a small sample size. 


*Create a contour plot of the sample size and population dispersion power simulation.*


```{r, echo = FALSE, results = FALSE, fig.show = 'hide'}
alpha = 0.05

# Start with a small number
n_sims = 10
p_vals = numeric(n_sims) 

# What was the observed standard deviation?
sd_obs

# specify the number of different standard deviation values to simulate:
# Start with a small number
n_sds = 100
pop_sds = seq(from = 0.05, to = 2.0, length.out = n_sds)

pop_sd_power = numeric(n_sds)

sample_sizes = seq(5, 100)

sim_output_3 = matrix(nrow = length(pop_sds), ncol = length(sample_sizes))

for(k in 1:length(pop_sds))
{
  pop_sd_k = pop_sds[k]
  
  for(j in 1:length(sample_sizes))
  {
    x_vals = seq(0, 100, length.out = sample_sizes[j])
    
    for(i in 1:n_sims)
    {
      fit_sim = linear_sim_fit(
        x = x_vals, 
        y = int_obs, 
        slope = slope_obs, 
        st_dev = pop_sd_k)
      p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
    }
    
    sim_output_3[k, j] = sum(p_vals < alpha) / n_sims
  }
  print(paste0("Testing standard deviation ", k, " of ", n_sds))
}

image(sim_output_3)

sim_3_dat = 
  list(
    power       = sim_output_3,
    sample_size = sample_sizes,
    pop_sd      = pop_sds)


# You should save your simulation results so you don't have to run it every time.
save(
  sim_3_dat, 
  file = here::here("data", "lab_ll_sim_output_dispersion_n_1000.RData"))
```


# Q3

*Include a figure of your contour plot in your report.*

```{r, echo = FALSE, warning = FALSE}
contour(
  x = sim_3_dat$pop_sd, y = sim_3_dat$sample_size, z = sim_3_dat$power,
  xlab = "standard deviation",
  ylab = "sample size",
  main = "Contour Plot of Statistical Power", 
  levels = seq(0, 1, length.out = 9),
  drawables = TRUE, 
  method = "edge")
```


# Q4

*Qualitatively describe the patterns you see in the contour plot. Make sure you discuss the effects of sample size and population dispersion on statistical power.*

# Q5



