---
title: "code_chunk_demo"
author: "Mandy Klehr"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tabset Example {.tabset} 
Use the curly brace and .tabset to create a tabset. 

## Tab 1

Tab 1 contents, code chunks

```{r load_penguin_data, eval=TRUE, echo=TRUE}

require(palmerpenguins)
head(penguins)

```

## Tab 2

Tab 2 contents, simple pair plot with penguins data

```{r simple_pairplot}
plot(penguins[, 3:5])
```

## Tab 3 

Tab 3 contents, hide code and change aspect of the simple pair plot

```{r simple_pairplot_change aspect ratio, echo = FALSE, fig.asp=1/1.5, fig.width=4}

plot(penguins[, 3:5])

#aspect ratio is greater than one, get a tall plot. Aspect ratio is less than one it will be a wide plot. 
```

# Next section

new code that is non-tabbed

# Different Penguin Plots {.tabset .tabset-pills}

The following tabs contain different plot types

## Histogram

```{r, fig.asp = 1/1.5}
hist(penguins$flipper_length_mm)
```


## Another Histogram

```{r, echo = FALSE}
hist(penguins$bill_length_mm)
```



