---
title: "DataCamp Assignment: R Basics"
author: Mandy Klehr
date: 9/10/2021
output: 
  html_document:
    theme: spacelab 
    toc: TRUE
    toc_float: TRUE
---

# Section 1: Variables

```{r}
a = "Mandy"
b1 = 45.6
b2 = "45.6"
c1 = c(0, 1, 2, 3)

b1 + c1
```

### Questions 

1. What type of data is contained in the variable a?
  - character

2.	What type of data is contained in the variable b1?
  - numeric

3.	What type of data is contained in the variable b2?
  - character

4.	What happens when you add b1 and b2 and why?
 -  R generates an error message because that operation is trying to add a numeric variable to a character, as is described in the error message: “non-numeric argument to binary operator” 

5.	Are the variables b1 and c1 of the same type?
  - Yes, they are both numeric. However, b1 is a single number, and c1 is a string of numbers.

6.	Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.
  - A string of numbers are returned because b1 as a single numeric (45.6) is added to all of the numbers within c1 as a string, thus 4 numbers are returned = [1] 45.6 46.6 47.6 48.6 

# Section 2: Vectors

### Questions

7. R code for *v1* below:
```{r}
v1 = c(-2:2)
v1
```

8. R code for *v2* below: 

```{r}
v2 = v1 * 3
v2
```

9. R code for *sum of v2* below:
```{r}
sum(v2)
```

# Section 3: Matrices

### Questions

10. R code for *mat_1* below:

```{r}
vec_4 = (1:12)
mat_1 = matrix(vec_4, nrow = 3,
              ncol = 4, byrow = TRUE)
mat_1
```

11. R code for *mat_2* below: 

```{r}
mat_2 = matrix(vec_4, nrow = 3, 
               ncol = 4, byrow = FALSE)
mat_2
```

# Section 3: Lists

### Questions

12. R code for creating *my_list_1* below:

```{r}
my_list_1 = list(5.2, "five point two", 0:5)
names(my_list_1) = c("two", "one", "three")
my_list_1
```

13. R code to select the *third element* in my_list_1 below: 

```{r}
my_list_1[3]
```

14. R code to select element named *"one"* in my_list_1 below: 

```{r}
my_list_1$one
```

# Section 3: Logical Tests and Subsetting Questions

### Questions

15. R code to create *my_bool_vec* below: 

```{r}
my_vec = rep(1:3, 5)
my_vec

my_bool_vec = my_vec == 3
data.frame(my_vec, my_bool_vec)
```

16. R code to perform subset of selecting elements of my_vec with value of 3 below:

```{r}
my_vec[my_bool_vec[TRUE]]
```

