---
title: 'Final Project: R Reference'
author: "Mandy Klehr"
date: "12/2/2021"
output: html_document
---

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

### here() 

The function here() can be used to locate data files in your directory. 

Both library() and require() can be used to load packages that are included or installed in R. Below I will show how to use library and require to load the *palmerpenguins* data. 

```{r}
library(palmerpenguins)

require(palmerpenguins)

```


I can use here to load the 

```{r}


```


Use these to show how to load the here and palmerpenguins packages - libary() and require()

Ginkgo data: use the 2021 ginkgo data to create a data.frame called ginkgo using:

here()
read.csv()


## Data Structures

### c() 

The function c() *combines* or *concatenates* its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
* I can’t combine character and numeric types in the same call to c()

Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

```


I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)

```

### length()

Factor a vecotr or a list. 

Using the example from the *c()* function we can figure out the length of the char_vec list: 
```{r}
## Use length to see how many ?items? are in the char_vec list:

length(char_vec)

```



## Subsetting




## Numerical Data Exploration


## Graphical Data Exploration


## Distribution Functions

