---
title: 'Lab 09: Modeling Data 2'
author: "Mandy Klehr -- I worked with John, Matt, Heather, and Bonnie"
date: "11/21/2021"
output:  
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes

---

```{r walkthrough, echo = FALSE, message = FALSE, results = FALSE, fig.show= 'hide', warning = FALSE}

require(here)
catrate = read.csv(here("data", "catrate.csv"))
head(catrate)

n_success = sum(catrate$success)
n_years = sum(catrate$years)
binom.test(n_success, n_years)

late_fill_rate = 2/7
normal_fill_rate = 1 - late_fill_rate

binom.test(
  n_success, 
  n_years,
  p = normal_fill_rate)

binom.test(
  n_success, 
  n_years, 
  p = normal_fill_rate, 
  alternative = "less")

veg = read.csv(here("data", "vegdata.csv"))
head(veg)

boxplot(pine ~ treatment, data = veg)

var.test(
  pine ~ treatment, 
  data = veg, 
  subset = treatment %in% c("control", "clipped"))


shapiro.test(veg$pine[veg$treatment=="control"])

shapiro.test(veg$pine[veg$treatment == "clipped"])

fligner.test(
  pine ~ treatment, 
  data = veg, 
  subset = treatment %in% c("control", "clipped"))

bartlett.test(pine ~ treatment, data = veg)

fligner.test(pine ~ treatment, data = veg)

t.test(
  pine ~ treatment, 
  data = veg, 
  subset = treatment %in% c("control", "clipped"))

wilcox.test(
  pine ~ treatment, 
  data = veg, 
  subset = treatment %in% c("control", "clipped"))

control = veg$pine[veg$treatment=="control"]
clipped = veg$pine[veg$treatment=="clipped"]

t.test(control, clipped, paired = TRUE)

wilcox.test(control, clipped, paired = TRUE)


#fala = penguins ~ sex * species
#fala = penguins ~ species * sex
#names = gsub("\\.", "\n", bxplt$names)
#las = 2
#xlab = ""

disp = read.csv(here("data", "dispersal.csv"))

disp

plot(disp$disp.rate.ftb, disp$disp.rate.eb)

cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use = "complete.obs")

cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use = "complete.obs",
  method = "spearman")



plot(
  ecdf(disp$disp.rate.ftb),
  verticals = TRUE)


```

```{r, echo = FALSE, results= FALSE, fig.show= 'hide', warning=FALSE}
plot(
  ecdf(disp$disp.rate.ftb),
  verticals = TRUE)

plot(
  ecdf(disp$disp.rate.eb),
  verticals = TRUE,
  lty = 3,
  add = TRUE)

ks.test(disp$disp.rate.ftb, disp$disp.rate.eb)

prop.test(c(4,16), c(40,250))

owls = matrix(c(16, 9, 4, 11), nrow = 2)
rownames(owls) = c("present", "absent")
colnames(owls) = c("old", "young")
chisq.test(owls)
owls

fisher.test(owls)

birds = read.csv(here("data", "bird.sta.csv"))
hab = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by = c("basin", "sub", "sta"))

#Create a contingency table for edge/interior and brown creeper presence/absence

table(birdhab$s.edge, birdhab$BRCR > 0)

#set the presence to be in the first column

br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]

br_creeper_table

#flinger is same (non-parametric) is non-normal; any distribution that is defined by parameters any named distribution is parametric? 
```

# Q1

There is no difference between brown creeper presence/absence in edge and interior habitats. In other words, brown creeper presence/absence in edge habitats is the same as brown creeper presence/absence in interior habitats because brown creeper presence/absence is not dependent on habitat type.  

```{r}
require(here)
birds = read.csv(here("data", "bird.sta.csv"))
hab = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by = c("basin", "sub", "sta"))

br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]
br_creeper_table

chisq.test(br_creeper_table)

```

# Q2

The p-value is well below 0.05 (1.386e-06) which indicates that we can reject the null hypothesis  and it is likely that brown creeper shows a significant habitat preference for interior habitats compared to edge habitats. This may not further signify this, because it is just the samples exhibited in our daa, but in our contingency table we see that at 29 of 173 plots in edge habitats the species was marked as present, whereas it was recorded at 214 of 773 plots in interior habitats.  


# Q3

```{r}
require(palmerpenguins)

fit_species = 
  lm(
    formula = body_mass_g ~ species, 
    data = penguins)
```

# Q4

```{r}
fit_sex = 
  lm(
    formula = body_mass_g ~ sex,
    data = penguins)

```

# Q5

```{r}
fit_both = 
  lm(
    formula = body_mass_g ~ sex * species, 
    data = penguins)
```

# Q6

```{r}
boxplot(body_mass_g ~ species, 
    data = penguins, 
    main = "Body mass of penguins \n by species",
    ylab = "Body mass (g)",
    xlab = '')
```

# Q7 

```{r}
boxplot(body_mass_g ~ sex, 
    data = penguins, 
    main = "Body mass of penguins \n by sex",
    ylab = "Body mass (g)",
    xlab = '')
```


# Q8

```{r}
bxplt = boxplot(body_mass_g ~ sex * species,
                data = penguins,
                plot = FALSE)


boxplot(body_mass_g ~ sex * species, 
    data = penguins, 
    main = "Body mass of penguins \n by sex and species",
    ylab = "Body mass (g)",
    names = gsub("\\.", "\n", bxplt$names),
    las = 2,
    xlab = '')

```


# Q9 

Based on visual interpretation alone, it looks like the body mass by species box plot may have problems fulfilling the heterogeneity assumption, because the variance (e.g., the size of the boxplot from the 1st to 3rd quartile) is small. However, this is just an indication of the variance and homogeneity or heterogeneity of the variance should be confirmed with a statistic test (as we do with the Bartlett's test below). 


# Q10 

The null hypothesis of Bartlett's test is that the variances in body mass for the different species of penguins or sex of penguin species are equal, e.g., all variances are homogeneous.  


# Q11

0.0500

```{r}
bartlett.test(body_mass_g ~ species, data = penguins)
```


# Q12

0.0319

```{r}
bartlett.test(body_mass_g ~ sex, data = penguins)
```


# Q13

0.1741

```{r}

dat_bm_sex = aggregate(
  body_mass_g ~ sex * species, 
  data = penguins, FUN = c)

bartlett.test(dat_bm_sex$body_mass_g)
```


# Q14

There appears to be an issue with heterogeneity of variances in the Bartlett test for the observations grouped by both sex and species as the p-value was 0.1741, so we could say that we would not assume that the variances in body mass between sex and species of penguins are equal. The other two Bartlett tests were below a p-value of 0.05 (marginally for species in Q11) and we can reject the null hypothesis, and assume that that there is homogeneity in variances of body mass and species, and body mass and sex when they are not grouped together. 

